Paper_Figure

### 2024-09-26 LMD-paper/paper_figure.Rmd
figure_change

---
title: "Paper_Figure"
author: "Ruiqi Li"
output: pdf_document
date: '2023-10-16'
---
# Load Data & Preprocess
```{r}
library("Seurat")
library("RColorBrewer")
library("dplyr")
library("patchwork")
library("ggplot2")
library("ggsignif")
library("LocalizedMarkerDetector")
source("paper_figure_function.R", echo = F)
source("paths.R")
figure_path = file.path(dir.path0,"figures")
dir.create(figure_path,recursive = TRUE)
setwd(figure_path)
```

@@

ggsave(filename = file.path(figure_path,"fig-known_marker_comparison.png"), plot = fig, width = 12, height = 6)
```
## Table
```{r}
library("reshape2")
wide_data = df_auc; wide_data$AUROC = round(wide_data$AUROC,3)
wide_data <- reshape2::dcast(wide_data, gt_set + Tissue + data_source ~ Method, value.var = "AUROC")
colnames(wide_data)[1:3] = c("Criterion","Tissue","DataSource")
wide_data[,1:3] = wide_data[,3:1]; colnames(wide_data)[1:3]=colnames(wide_data)[3:1]
wide_data <- wide_data[order(wide_data$DataSource,wide_data$Tissue,wide_data$Criterion),]
wide_data[,1:3] = apply(wide_data[,1:3],2,function(x) gsub("[\n]|Criterion1-|Criterion2-","",x))
wide_data[,1:2] = apply(wide_data[,1:2],2,function(x) {x[duplicated(x)] = "";x})
wide_data$Tissue[17] = "Motor Cortex(Mouse)"
wide_data$Tissue[13] = "Motor Cortex(Human)"
wide_data$Tissue = gsub("Human-|Mouse-","",wide_data$Tissue)
# calculate average
avg_tmp = wide_data %>%
    group_by(Criterion) %>%
    summarize(across(3:(ncol(wide_data)-1), ~ round(mean(.x, na.rm = TRUE), 3)))
avg_tmp$DataSource = c("Average",""); avg_tmp$Tissue = "-"
wide_data = rbind(wide_data,avg_tmp)
bold_max <- function(vec) {
  max_val <- max(vec, na.rm = TRUE)
  vec = unlist(ifelse(vec == max_val, paste0("\\textbf{", vec, "}"), vec))
  return(vec)
}
wide_data[,4:ncol(wide_data)] = t(apply(wide_data[,4:ncol(wide_data)],1,bold_max))
wide_data[is.na(wide_data)] = "-"

write.csv(wide_data, file = file.path(folder.path,"AUROC_table.csv"), row.names = FALSE)
```
# SupFig-RunTime
```{r}
df_runtime = do.call(rbind,
                lapply(1:length(tissue_ls),function(i){
                  df_all = NULL
                  for(j in tissue_ls[[i]]){
                    df = read.table(file.path(folder.path, names(tissue_ls)[i],paste0(j,"_runtime.csv")),header = TRUE)
                    df$'Tissue' = j
                    df_all = rbind(df_all,df)
                  }
                  df_all$'data_source' = names(tissue_ls)[i]
                  df_all}))
df_runtime$Method = method_fullname[df_runtime$Method]
df_runtime$Tissue = tissue_full_name[df_runtime$Tissue]
df_runtime$data_source = data_source_full_name[df_runtime$data_source]
```
## Figure
```{r}
coldef = setNames(
  colorRampPalette(brewer.pal(8, "Set1"))(length(method_fullname)),
  method_fullname)

@@

# Fig - Gene Ranking Comparisons between LMD and Other Methods
```{r}
dir.path <- dir.path0
data_source = "tabular_muris"
tissue_name = "marrow_facs"
df_benchmark = read.table(file.path(dir.path,"benchmark", data_source, paste0(tissue_name,"_benchmark_rank_table.csv")),header = TRUE)
rownames(df_benchmark) = df_benchmark$gene
tiss = readRDS(file.path(dir.path,data_source, paste0(tissue_name,".rds")))
method_vs = "lmd"
method_vs_set = c("wilcox_no_filter","hotspot","marcopolo","semi","haystack")
```
```{r}
sub = do.call(c,lapply(method_vs_set,function(method){
  sub = c(which(df_benchmark[,method] <= 100 & df_benchmark[,method_vs] > 500),
          which(df_benchmark[,method] > 500 & df_benchmark[,method_vs] <= 100) )
  sub
}))
df_label_all = df_benchmark[unique(sub),c(method_vs,method_vs_set)]
# df_label_all$'legend' = sapply(seq(nrow(df_label_all)),function(i){
#   paste(paste(method_fullname[colnames(df_label_all)],df_label_all[i,],sep = ":"),collapse = "\n")
# })
g = c("Cd19","Naaa","Mogat2","Grina","Abca13","Zf12","Il2rb","Phgdh","Vpreb2","Fyb")
names(g) = rep(method_vs_set,each = 2)
pl = lapply(1:length(g),function(i){
  x = g[i]
  subtitle = sprintf("%s:%d, %s:%d",method_fullname[method_vs],df_label_all[x,method_vs],method_fullname[names(g)[i]],df_label_all[x,names(g)[i]])
  color_label = ifelse(df_label_all[x,method_vs] > df_label_all[x,names(g)[i]],"blue","red")
  FeaturePlot(tiss,features = x,order =TRUE) + labs(color = "Expression") + labs(title = x,subtitle = subtitle) + 
    theme(
      plot.title = element_text(face="bold", size = 30,color = color_label),
      plot.subtitle = element_text(size = 20),
      legend.title = element_text(size = 20),
      legend.text = element_text(size = 15),
      axis.title = element_blank(),
      axis.text = element_blank(),
      axis.ticks=element_blank(),
      axis.line = element_blank())
  # + annotation_custom(grob = grid::textGrob(df_label_all[x,"legend"], x = 1, y = 1, just = c("right", "top"),gp = grid::gpar(col = "red", fontsize = 10)))
})
pl = pl[order(seq(length(g))%%2,decreasing = TRUE)]
pl[[1]] = pl[[1]] + labs(y = "Favored by LMD") + theme(axis.title.y = element_text(size = 20,face="bold", colour = "red"))
pl[[(length(pl)/2)+1]] = pl[[(length(pl)/2)+1]] + labs(y = "Favored by Other methods") + theme(axis.title.y = element_text(size = 20,face="bold", colour = "blue"))

@@

# SupFig - Density Index
```{r}
# Set the color code for all methods
method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_no_filter","hvg")
method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v4","HVGs"),method_ls)
coldef = setNames(
  colorRampPalette(brewer.pal(8, "Set1"))(length(method_fullname)),
  method_fullname)
```

```{r}
data_source = "tabular_muris"
dir.path <- dir.path0
pl = lapply(c("marrow_facs","lung_facs","pancreas_facs"),function(tissue_name){
score_measure = read.table(file.path(dir.path,"benchmark",data_source,paste0(tissue_name,"_DI.csv")),header = TRUE)
score_measure$Method = ifelse(score_measure$Method %in% names(method_fullname),method_fullname[score_measure$Method],score_measure$Method)

@@

# Only keep Bone Marrow
fig = wrap_plots(pl[[1]],ncol = 1) + plot_layout(guides = "collect") + plot_annotation(tag_levels = "A") & theme(legend.position = 'bottom', plot.tag = element_text(size = 20))
ggsave(filename = file.path(figure_path,"supfig-densityindex.png"), plot = fig, width = 12, height = 5)
```

@@

>>>>

---
title: "Paper_Figure"
author: "Ruiqi Li"
output: pdf_document
date: '2023-10-16'
---
# Load Data & Preprocess
```{r}
library("Seurat")
library("RColorBrewer")
library("dplyr")
library("patchwork")
library("ggplot2")
# library("ggsignif")
library("LocalizedMarkerDetector")
source("paper_figure_function.R", echo = F)
source("paths.R")
figure_path = file.path(dir.path0,"figures")
dir.create(figure_path,recursive = TRUE)
setwd(figure_path)
```

@@

ggsave(filename = file.path(figure_path,"fig-known_marker_comparison.png"), plot = fig, width = 12, height = 6)
```
## Table
```{r}
library("reshape2")
wide_data = df_auc; 
# wide_data$AUROC = round(wide_data$AUROC,3)
wide_data <- reshape2::dcast(wide_data, gt_set + Tissue + data_source ~ Method, value.var = "AUROC")
colnames(wide_data)[1:3] = c("Criterion","Tissue","DataSource")
wide_data[,1:3] = wide_data[,3:1]; colnames(wide_data)[1:3]=colnames(wide_data)[3:1]
wide_data <- wide_data[order(wide_data$DataSource,wide_data$Tissue,wide_data$Criterion),]
wide_data[,1:3] = apply(wide_data[,1:3],2,function(x) gsub("[\n]|Criterion1-|Criterion2-","",x))
wide_data[,1:2] = apply(wide_data[,1:2],2,function(x) {x[duplicated(x)] = "";x})
wide_data$Tissue[which(wide_data$Tissue == "Mouse-Motor Cortex")] = "Motor Cortex(Mouse)"
wide_data$Tissue[which(wide_data$Tissue == "Human-Motor Cortex")] = "Motor Cortex(Human)"
wide_data$Tissue = gsub("Human-|Mouse-","",wide_data$Tissue)

# calculate median rank
ranked_data <- data.frame(t(apply(wide_data[,4:ncol(wide_data)],1,function(x) rank(-x, na.last = "keep"))), check.names = FALSE)
ranked_data$Criterion = wide_data$Criterion
median_rank = ranked_data %>% group_by(Criterion) %>% summarize(across(1:(ncol(ranked_data)-1), ~ median(.x, na.rm = TRUE)))
median_rank$DataSource = c("Median Rank",""); median_rank$Tissue = "-"
wide_data = rbind(wide_data,median_rank)

# # calculate average
# avg_tmp = wide_data %>%
#     group_by(Criterion) %>%
#     summarize(across(3:(ncol(wide_data)-1), ~ mean(.x, na.rm = TRUE)))
# avg_tmp$DataSource = c("Average",""); avg_tmp$Tissue = "-"
# wide_data = rbind(wide_data,avg_tmp)

wide_data[,4:ncol(wide_data)] = round(wide_data[,4:ncol(wide_data)],3)

bold_max <- function(vec) {
  max_val <- max(vec, na.rm = TRUE)
  vec = unlist(ifelse(vec == max_val, paste0("\\textbf{", vec, "}"), vec))
  return(vec)
}
wide_data[,4:ncol(wide_data)] = t(apply(wide_data[,4:ncol(wide_data)],1,bold_max))
wide_data[is.na(wide_data)] = "-"

# write.csv(wide_data, file = file.path(folder.path,"AUROC_table.csv"), row.names = FALSE)
```
## Colored Table figure
```{r}
library(tidyr)

# separate wide_data into two by true_set
df_head = wide_data %>% filter(Criterion == " CellMarkerDB") %>% select(c("DataSource","Tissue"))
df_head[nrow(df_head),"DataSource"] = ""
df_head[nrow(df_head),"Tissue"] = "Median Rank"
df_head$DataSource = c(
  rep(df_head$DataSource[1],3),
  rep(df_head$DataSource[4],6),
  df_head$DataSource[10])

pl = lapply(1:2,function(i){
trueset = unique(wide_data$Criterion)[i]
wide_data_sub = wide_data %>% filter(Criterion == trueset) %>% select(all_of(unname(method_fullname))) %>% mutate(across(everything(), ~ as.numeric(gsub("\\\\textbf\\{(.*)\\}", "\\1", .)))) %>% mutate(df_head)

ranked_data_sub = ranked_data %>% filter(Criterion == trueset) %>% select(all_of(unname(method_fullname))) %>% mutate(df_head[1:(nrow(df_head)-1),])
ranked_data_sub = rbind(ranked_data_sub,wide_data_sub[nrow(wide_data_sub),])

# Melt the data into long format for ggplot2
df_long <- wide_data_sub %>%
  pivot_longer(cols = all_of(unname(method_fullname)), names_to = "Method", values_to = "AUROC") %>%
  left_join(ranked_data_sub %>%
              pivot_longer(cols = all_of(unname(method_fullname)), names_to = "Method", values_to = "Rank"),
            by = c("DataSource","Tissue","Method"))

df_long$Method = factor(df_long$Method, levels = rev(colnames(wide_data)[4:10]))
df_long$DataSource = factor(df_long$DataSource, levels = unique(df_long$DataSource))

# Plot the heatmap
p = ggplot(df_long, aes(x = Tissue, y = Method, fill = Rank)) +
    geom_tile(color = "white") + 
    scale_fill_distiller(palette = "YlOrBr", direction = -1, na.value = "white") +
    # scale_fill_gradient(low = "brown", high = "white", na.value = "white") +
    geom_text(aes(label = AUROC), size = 3) + 
    theme_minimal() + facet_grid(~ DataSource, scales = "free",space = "free", switch = "x") + ggtitle(paste0("True markerset ",i)) + 
    theme(plot.title = element_text(size = 15,face = "bold"),
          axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
          strip.placement = 'outside',
          axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          panel.grid = element_blank(), 
          strip.text = element_text(size = 10,face = "bold"),
          axis.text.y = element_text(face = "bold", size = 10))

if(i==2){p = p + theme(axis.text.y = element_blank())}

p
})

p = wrap_plots(pl, nrow = 1) + plot_layout(guides = "collect")

ggsave(filename = file.path(figure_path,"fig-known_marker_comparison.png"), plot = p, width = 12, height = 6)
```

# SupFig-RunTime
```{r}
df_runtime = do.call(rbind,
                lapply(1:length(tissue_ls),function(i){
                  df_all = NULL
                  for(j in tissue_ls[[i]]){
                    df = read.table(file.path(folder.path, names(tissue_ls)[i],paste0(j,"_runtime.csv")),header = TRUE)
                    df$'Tissue' = j
                    df_all = rbind(df_all,df)
                  }
                  df_all$'data_source' = names(tissue_ls)[i]
                  df_all}))
df_runtime$Method = method_fullname[df_runtime$Method]
df_runtime$Tissue = tissue_full_name[df_runtime$Tissue]
df_runtime$data_source = data_source_full_name[df_runtime$data_source]
```
## Figure
```{r}
coldef = setNames(
  colorRampPalette(brewer.pal(8, "Set1"))(length(method_fullname)),
  method_fullname)

@@

# Fig - Gene Ranking Comparisons between LMD and Other Methods
```{r}
dir.path <- dir.path0
data_source = "tabular_muris"
tissue_name = "marrow_facs"
df_benchmark = read.table(file.path(dir.path,"benchmark", data_source, paste0(tissue_name,"_benchmark_rank_table.csv")),header = TRUE)
rownames(df_benchmark) = df_benchmark$gene
tiss = readRDS(file.path(dir.path,data_source, paste0(tissue_name,".rds")))
method_vs = "lmd"
method_vs_set = c("wilcox_no_filter","hotspot","marcopolo","semi","haystack")
```
```{r}
sub = do.call(c,lapply(method_vs_set,function(method){
  sub = c(which(df_benchmark[,method] <= 100 & df_benchmark[,method_vs] > 500),
          which(df_benchmark[,method] > 500 & df_benchmark[,method_vs] <= 100) )
  sub
}))
df_label_all = df_benchmark[unique(sub),c(method_vs,method_vs_set)]
# df_label_all$'legend' = sapply(seq(nrow(df_label_all)),function(i){
#   paste(paste(method_fullname[colnames(df_label_all)],df_label_all[i,],sep = ":"),collapse = "\n")
# })
g = c("Cd19","Naaa","Mogat2","Grina","Abca13","Zf12","Il2rb","Phgdh","Vpreb2","Fyb") # <100&>500
names(g) = rep(method_vs_set,each = 2)
pl = lapply(1:length(g),function(i){
  x = g[i]
  subtitle = sprintf("%s:%d, %s:%d",method_fullname[method_vs],df_label_all[x,method_vs],method_fullname[names(g)[i]],df_label_all[x,names(g)[i]])
  color_label = ifelse(df_label_all[x,method_vs] > df_label_all[x,names(g)[i]],"blue","red")
  FeaturePlot(tiss,features = x,order =TRUE) + labs(color = "Expression") + labs(title = x,subtitle = subtitle) + 
    theme(
      plot.title = element_text(face="bold", size = 30,color = color_label),
      plot.subtitle = element_text(size = 20),
      legend.title = element_text(size = 20),
      legend.text = element_text(size = 15),
      axis.title = element_blank(),
      axis.text = element_blank(),
      axis.ticks=element_blank(),
      axis.line = element_blank())
  # + annotation_custom(grob = grid::textGrob(df_label_all[x,"legend"], x = 1, y = 1, just = c("right", "top"),gp = grid::gpar(col = "red", fontsize = 10)))
})
pl = pl[order(seq(length(g))%%2,decreasing = TRUE)]
pl[[1]] = pl[[1]] + labs(y = "Favored by LMD") + theme(axis.title.y = element_text(size = 20,face="bold", colour = "red"))
pl[[(length(pl)/2)+1]] = pl[[(length(pl)/2)+1]] + labs(y = "Favored by Other methods") + theme(axis.title.y = element_text(size = 20,face="bold", colour = "blue"))

@@

# SupFig - Density Index
```{r}
# Set the color code for all methods
method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_no_filter","hvg")
method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v4","HVGs"),method_ls)
coldef = setNames(
  colorRampPalette(brewer.pal(8, "Set1"))(length(method_fullname)),
  method_fullname)
```
## Curve Plot & Coverage rate
```{r}
data_source = "tabular_muris"
dir.path <- dir.path0
pl = lapply(c("marrow_facs","lung_facs","pancreas_facs"),function(tissue_name){
score_measure = read.table(file.path(dir.path,"benchmark",data_source,paste0(tissue_name,"_DI.csv")),header = TRUE)
score_measure$Method = ifelse(score_measure$Method %in% names(method_fullname),method_fullname[score_measure$Method],score_measure$Method)

@@

# Only keep Bone Marrow
fig = wrap_plots(pl[[1]],ncol = 1) + plot_layout(guides = "collect") + plot_annotation(tag_levels = "A") & theme(legend.position = 'bottom', plot.tag = element_text(size = 20))
ggsave(filename = file.path(figure_path,"supfig-densityindex.png"), plot = fig, width = 12, height = 5)
```
## Colored Table
```{r}
data_source = "tabular_muris"
dir.path <- dir.path0
# tissue_name = "marrow_facs" # marrow_facs_granulocyte

pl = lapply(c("marrow_facs","marrow_facs_granulocyte"),function(tissue_name){
score_measure = read.table(file.path(dir.path,"benchmark",data_source,paste0(tissue_name,"_DI.csv")),header = TRUE)
score_measure$Method = ifelse(score_measure$Method %in% names(method_fullname),method_fullname[score_measure$Method],score_measure$Method)
total_gene = score_measure %>% filter(Method == "All genes") %>%.$TopGene
score_measure = score_measure %>%
  filter(Method == "All genes") %>% 
  slice(rep(1:n(), each = length(method_fullname))) %>% 
  mutate(Method = method_fullname) %>%
  bind_rows(score_measure) %>% filter(Method != "All genes") %>% filter(TopGene %in% c(100,300,500,1000,3000,5000,10000,total_gene)) %>% mutate(TopGene = as.factor(TopGene)) %>% mutate(Method = factor(Method,levels = rev(method_fullname)))

levels(score_measure$TopGene)[nlevels(score_measure$TopGene)] = sprintf("All(%d)",total_gene)

fig = ggplot(score_measure, aes(x = TopGene, y = Method, fill = DensityIndex)) + 
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white",high = "red") +
  geom_text(aes(label = round(DensityIndex,3)), size = 3) + theme_minimal() + xlab("# of Top Genes") + ylab("Method") + 
  ggtitle(ifelse(tissue_name == "marrow_facs", "Bone Marrow","Bone Marrow (granulocyte)")) + 
  theme(plot.title = element_text(size = 20, face = "bold"),
          axis.text = element_text(size = 10),
          axis.title = element_text(size = 10),
          panel.grid = element_blank(),
        axis.ticks = element_line())
fig
})

fig = wrap_plots(pl,ncol = 1) + plot_annotation(tag_levels = "A") & theme(plot.tag = element_text(size = 20))
ggsave(filename = file.path(figure_path,"supfig-densityindex.png"), plot = fig, width = 10, height = 10)
```

@@

# SupFig - sample-specificity for each module
```{r}
df = read.csv(file.path(dir.path,"intermediate_data","cell_prop_in_module_table.csv"))
df$gene_props = df$gene_props * 100
df = df %>% filter(module=="Module12" | sample!="E14.5_WT") %>% filter(gene_props > 5 & gene_props < 50)
p = ggplot(df,
       aes(x=gene_props, y=cell_props, linetype = sample,color = sample)) +
  geom_line() + facet_wrap(~module,scales = "free") +
  labs(x = "% of genes", y = "Proportion of cells", title = "Expression of Module 12 genes in DC samples") + 
  theme(legend.position="right",
          plot.title = element_text(size = 10),
          axis.title = element_text(size = 10),
          axis.text = element_text(size = 5),
          legend.title = element_text(size = 5),
          legend.text = element_text(size = 5))
ggsave(filename = file.path(figure_path,"supfig-smom2_sample_specificity.png"), plot = p, width = 4, height = 3)
```
